---
title: "Introduction to Probability"
author: "William Murrah"
csl: apa.csl
date: "10/17/2014"
bibliography: introProb.bib
output: 
   pdf_document:
      includes:
         in_header: header.tex
---

```{r opts}
library(knitr)
opts_knit$set(root.dir='../../')
opts_chunk$set(comment=NA, prompt=TRUE)
```

```{r data}

```

# Notation and Symbols

This uses [@lynch2007introduction]. And [@ipsur2013].

# What is Probability?

There are at least two views about what probability is. 

## An adequate set of operations 


### Conjunction

The logical product, or conjunction, indicates that both $A$ and $B$ are true, and is expressed as:

$$AB.$$

In R conjunction can be coded with the logical *and* operator (`&`):
```{r conjunction, eval=FALSE}

A & B 

```

### Disjunction

The logical sum, or disjunction, indicates that at least one of $A$ or $B$ is true, and is expressed as:

$$A + B.$$

In R, disjunction is coded with the logical *or* operator:
```{r disjunction, eval=FALSE}

A | B
```

### Negation

The logical denial, or negation, of a proposition is expressed with a bar over the symbol. For example, the negation of $A$ is expressed as:

$$\bar{A}$$

which indicates that proposition $A$ is false.

In R, logical negation is coded using the exclamation point in front of the object:
```{r negation, eval=FALSE}

!A
```

## Some trivial basic identities

First we initialize the truth of $A$, $B$ and $C$ in R:

```{r initAB}
A <- TRUE
B <- TRUE
C <- TRUE
```

### Idempotence

$AA = A$ 

$A + A = A$
```{r idempotence}
(A&A )== A
(A | A) == A
```

### Commutativity

$AB = BA$

$A + B = B + A$

```{r commitativity}
(A&B) ==( B&A)
(A | B) ==( B | A)
```

### Associativity

$A(BC) = (AB)C = ABC$

$A + (B + C) = (A + B) + C = A + B + C$

```{r associativity}
(A&(B&C)) == ((A&B)&C == A&B&C)
(A | (B | C)) == ((A | B) | C == A | B | C)
```

### Distributivity

$A(B + C) = AB + AC$

$A + (BC) = (A + B)(A + C)$
```{r distributivity}
(A&(B | C)) == (A&B | A&C)
(A | (B&C)) == ((A | C)&(A | C))
```

### Duality

If $C = AB$, then $\bar{C} = \bar{A} + \bar{B}$

If $D = A + B$, then $\bar{D} = \bar{A}\bar{B}$
```{r duality}
C  <-( A&B) 
!C == (!A | !B)
D <- A | B
!D == (!A&!B) # NOTE: the parentheses are needed!
```

### Implication

The expression

$$A \implies {B}$$

can be read as '$A$ implies $B$'. This statement does not indicate that either proposition is true, but instead simply means that $A\bar{B}$ is false or equivalently, $(\bar{A} + B)$ is true. This can be written as a logical equation as:

$$A = AB,$$

and coded in R as:
```{r implies, eval=FALSE}

A == A&B
```

Discuss logical and numerical '='. 

# References